Sparse Function-space Representation \\ of Neural Networks

### 2023-05-15 paper/main.tex
rewrite section 3

\paragraph{Dual parameters from NN}
Given we are making a Laplace approximation of the neural network we can remove the expectation over the posterior to find the following approximate dual variables (see Ch.~3.4.1 in \cite{rasmussen2006gaussian} for derivation). Meaning we arrive at the following dual variables.
\begin{equation}
\label{eq:dual_param_laplace}
\hat{\alpha}_i \coloneqq \nabla_{f_i}\log p(y_i \mid f_i),  \quad \text{and} \quad
\hat{\beta}_i \coloneqq - \nabla^2_{f_i f_i}\log p(y_i \mid f_i).
\end{equation}

Replacing \cref{eq:dual_param_laplace} in to \cref{eq-laplace-approx-function-space} and we arrive at our GP model based on the converged neural network. Again this is similar to what was arrived at for \citet{immerImprovingPredictionsBayesian2021} for the posterior variance function, but they instead use the $f_{\vw^*}(\vx)$ for the posterior mean and fail to spot the significance of the dual variables. The problem with \cref{eq:gp_pred} is that to make predictions and compute variances we must incur a cost of $O(n^3)$ which limits the use of the GP on large data sets.

\paragraph{Sparsifying the NNGP}
\label{sec:sparse-dual-gp}
%
Sparse Gaussian processes are a way to reduce the computational complexity of a full GP using a low-rank approximation. Many sparsifying approaches exist, we refer the reader to \cite{quinonero2005unifying} for an overview. Given that we have our dual variables derived from our neural network predictions and kernel function, we could use any approximation. We decide to follow the sparsifying approach of \citet{titsias2009variational} and also used in the DTC approximaton \cite{quinonero2005unifying}, which is to define the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) q(\vu)  d\vu$. Given that we have $p(f_i  \mid \vu)$ determined by our GP prior, the goal is to find a $q(\vu)$.

As was shown in \cite{adam2021dual}, the posterior under this model has a particular structure similar to \cref{eq:gp_pred}. In the paper, they spot the dual variables' importance for approximate inference but write them in the natural parameter form as this form is more suitable for optimization through natural gradients. In order to make the link to the dual variables defined in \cref{eq:dual_param} we write them in the sparse GP in the dual variable form. 
%
\begin{equation} 
	\valpha_{\vu}  =  \sum_{i \in \mathcal{D}_{n}}  \vkzi \, \hat{\alpha}_{i}, \quad
	\vbeta_{\vu} =  \sum_{i \in \mathcal{D}_{n}} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
\label{eq:dual_sparse}
\end{equation}
the sparse dual variables are now a sum over all data points, with $\valpha_{\vu} \in \R^{M}$ and $\vbeta_{\vu} \in \R^{M  \times M}$. Using the sparse definition of the dual variables our sparse GP posterior takes the following form:
\begin{equation}\label{eq:dual_sparse_post}
   \myexpect_{q_{\vu}(\vf)}[f_i] = \vkzs^{\T} \MKzz^{-1}   \valpha_{\vu} , \quad 
   \textrm{Var}_{q_{\vu}(\vf)}[f_i]  = \kappa_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \vbeta_{\vu})^{-1} ]\vkzs
\end{equation}
where $\MKzz$ and $\vkzs$ are define similarly to $\MKxx$ and $\vk_{\vx i}$ but now over the $\MZ \in \R^{M \times D} $ inducing points instead of the full dataset $\MX$. The key quantities we need to make predictions from our sparse GP from the converged neural network are $(\hat{\alpha}_i, \hat{\beta}_i)$ and a kernel function $\kappa$. Contrasting \cref{eq:dual_sparse_post} and \cref{eq:gp_pred} we can see that the computational complexity is now $\mathcal{O}(m^3)$ to $\mathcal{O}(n^3)$ where $m \ll n$.  Crucially given the structure of our probabilistic model, our sparse dual variables \cref{eq:dual_sparse} are a compact form of the full model projected using the kernel. Our approach contrasts to \citet{immer2021improving}, which uses a subset of the full model, a less principled approach to build a sparse model given it ignores the information in the dataset. It is important to highlight that the sparsifying approach is independent of the approximate inference technique. The \cref{eq:dual_param} are determined from the neural network. More complicated inference techniques, such as variational inference, could be used, but given its simplicity, we only need a trained neural network; we use a Laplace approximation in our experiments.

>>>>

\paragraph{Dual parameters from NN}
Given we are making a Laplace approximation of the neural network we can remove the expectation over the posterior (see Ch.~3.4.1 in \cite{rasmussen2006gaussian} for derivation). Meaning we arrive at the following dual variables.
\begin{equation}
\label{eq:dual_param_laplace}
\hat{\alpha}_i \coloneqq \nabla_{f_i}\log p(y_i \mid f_i),  \quad \text{and} \quad
\hat{\beta}_i \coloneqq - \nabla^2_{f_i f_i}\log p(y_i \mid f_i).
\end{equation}

Substiuting \cref{eq:dual_param_laplace} in to \cref{eq:gp_pred} and we arrive at our GP model based on the converged neural network. Again this is similar to what was derived in \citet{immerImprovingPredictionsBayesian2021} for the posterior variance function, but they instead use the $f_{\vw^*}$ for the posterior mean and fail to spot the significance of the dual variables. The problem with \cref{eq:gp_pred} is that to make predictions and compute variances we must incur a cost of $O(n^3)$ which limits the use of the GP on large data sets.

\paragraph{Sparsifying the NNGP}
\label{sec:sparse-dual-gp}
%
Sparse Gaussian processes are a way to reduce the computational complexity of a full GP using a low-rank approximation. Many sparsifying approaches exist, we refer the reader to \cite{quinonero2005unifying} for an overview. Given that we have our dual variables derived from our neural network predictions and a kernel function, we could essentially use any approximation. We decide to follow the sparsifying approach of \citet{titsias2009variational} and also used in the DTC appeoximation \cite{quinonero2005unifying}, which is to define the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) q(\vu)  d\vu$. Given that we have $p(f_i  \mid \vu)$ determined by our GP prior, the goal is to find a $q(\vu)$.

As was shown in \cite{adam2021dual}, the posterior under this model has a particular structure similar to \cref{eq:gp_pred}. In the paper, they spot the dual variables' importance for approximate inference but write them in the natural parameter form as this form is more suitable for optimization through natural gradients. In order to make the link to the dual variables defined in \cref{eq:dual_param} we write them in the sparse GP in the dual variable form. 
%
\begin{equation} 
	\valpha_{\vu}  =  \sum_{i \in \mathcal{D}_{n}}  \vkzi \, \hat{\alpha}_{i}, \quad
	\vbeta_{\vu} =  \sum_{i \in \mathcal{D}_{n}} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
\label{eq:dual_sparse}
\end{equation}
the sparse dual variables are now a sum over all data points, with $\valpha_{\vu} \in \R^{M}$ and $\vbeta_{\vu} \in \R^{M  \times M}$. Using the sparse definition of the dual variables our sparse GP posterior takes the following form:
\begin{equation}\label{eq:dual_sparse_post}
   \myexpect_{q_{\vu}(\vf)}[f_i] = \vkzs^{\T} \MKzz^{-1}   \valpha_{\vu} , \quad 
   \textrm{Var}_{q_{\vu}(\vf)}[f_i]  = \kappa_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \vbeta_{\vu})^{-1} ]\vkzs
\end{equation}
where $\MKzz$ and $\vkzs$ are define similarly to $\MKxx$ and $\vk_{\vx i}$ but now over the $\MZ \in \R^{M \times D} $ inducing points instead of the full dataset $\MX$. The key quantities we need to make predictions from our sparse GP from the converged neural network are $(\hat{\alpha}_i, \hat{\beta}_i)$ and a kernel function $\kappa$. Contrasting \cref{eq:dual_sparse_post} and \cref{eq:gp_pred} we can see that the computational complexity is now $\mathcal{O}(m^3)$ to $\mathcal{O}(n^3)$ where $m \ll n$.  Crucially given the structure of our probabilistic model, our sparse dual variables \cref{eq:dual_sparse} are a compact form of the full model projected using the kernel.

Our approach contrasts to \citet{immer2021improving}, which uses a subset of the full model to try to reduce complexity, a less principled approach to build a sparse model given it ignores the information in the dataset. It is important to highlight that the sparsifying process is independent of the approximate inference technique. The \cref{eq:dual_param} are from the neural network definition \cref{eq-empirical-risk}. More complicated inference techniques, such as variational inference, could be used, but given its simplicity, we only need a trained neural network; we use a Laplace approximation in our experiments.
